---
title:     "Cirrincione T Study - Statistical Analysis"
updated:   1/24/24
author:    Rene coig
output:    html_notebook
---
Study Description: The purpose of this project is to determine how testosterone affects other medications and the natural bacteria that lives in the gut.

Notebook Description: Wilkoxon signed rank tests. 

R version 4.4.2 "Pile of Leaves"

Load in data
```{r}
rm(list = ls())#clearing environment
load("_output/NCA_Analysis.RData")
```
Test Tmax with paired Wilcoxon signed rank test.
```{r}
Test_tmax1_2<-PK_analysis_wide%>%
  select(Tmax_1, Tmax_2)%>%
  filter(!(is.na(Tmax_2))) #N = 11
Tmax_1 <- Test_tmax1_2$Tmax_1
Tmax_2 <- Test_tmax1_2$Tmax_2
Tmax_1
Tmax_2
Test_tmax1_2 <- wilcox.test(Tmax_1,Tmax_2,paired = TRUE)
Test_tmax1_2

Test_tmax1_3<-PK_analysis_wide%>%
  select(Tmax_1, Tmax_3) #N = 14
Tmax_1 <- Test_tmax1_3$Tmax_1
Tmax_3 <- Test_tmax1_3$Tmax_3
Tmax_1
Tmax_3
Test_tmax1_3 <- wilcox.test(Tmax_1,Tmax_3,paired = TRUE)
Test_tmax1_3
```
Test AUC (0-last) with paired Wilcoxon signed rank test.
```{r}
Test_AUC1_2<-PK_analysis_wide%>%
  select(AUClast_1, AUClast_2)%>%
  filter(!(is.na(AUClast_2))) #N = 11
AUClast_1 <- Test_AUC1_2$AUClast_1
AUClast_2 <- Test_AUC1_2$AUClast_2
AUClast_1
AUClast_2
Test_AUC1_2 <- wilcox.test(AUClast_1,AUClast_2,paired = TRUE)
Test_AUC1_2

Test_AUC1_3<-PK_analysis_wide%>%
  select(AUClast_1, AUClast_3) #N = 14
AUClast_1 <- Test_AUC1_3$AUClast_1
AUClast_3 <- Test_AUC1_3$AUClast_3
AUClast_1
AUClast_3
Test_AUC1_3 <- wilcox.test(AUClast_1,AUClast_3,paired = TRUE)
Test_AUC1_3
```
Test AUC (0-4) with paired Wilcoxon signed rank test.
```{r}
Test_AUC4_1_2<-PK_analysis_wide%>%
  select(AUC4_1, AUC4_2)%>%
  filter(!(is.na(AUC4_2))) #N = 11
AUC4_1 <- Test_AUC4_1_2$AUC4_1
AUC4_2 <- Test_AUC4_1_2$AUC4_2
AUC4_1
AUC4_2
Test_AUC4_1_2 <- wilcox.test(AUC4_1,AUC4_2,paired = TRUE)
Test_AUC4_1_2

Test_AUC4_1_3<-PK_analysis_wide%>%
  select(AUC4_1, AUC4_3) #N = 14
AUC4_1 <- Test_AUC4_1_3$AUC4_1
AUC4_3 <- Test_AUC4_1_3$AUC4_3
AUC4_1
AUC4_3
Test_AUC4_1_3 <- wilcox.test(AUC4_1,AUC4_3,paired = TRUE)
Test_AUC4_1_3
```